\documentclass[11pt, article, oneside]{memoir}

% Some useful packages
\usepackage
{
  microtype,
  blindtext,
  multirow,
  makecell,
  graphicx,
  hyperref,
  nicematrix,
  mathtools,
  amsthm,
  amssymb,
  thmtools,
  bm,
  siunitx,
  caption
}

%% PDF Metadata
\hypersetup
{
  hidelinks,
  pdftitle    = Title,
  pdfauthor   = Ashnah Khalid Khan
}


% Page Layout for Printing A4
\settrimmedsize{297mm}{210mm}{*}
\setlength{\trimtop}{0pt}
\setlength{\trimedge}{\stockwidth}
\addtolength{\trimedge}{-\paperwidth}
\settypeblocksize{634pt}{448.13pt}{*}
\setulmargins{4cm}{*}{*}
\setlrmargins{*}{*}{1}
\setmarginnotes{17pt}{51pt}{\onelineskip}
\setheadfoot{\onelineskip}{2\onelineskip}
\setheaderspaces{*}{2\onelineskip}{*}
\checkandfixthelayout%


% Page Style
\makepagestyle{SD}
\nouppercaseheads
\makeoddhead{SD}{\textsc{Theory of Automata}}{}{\textsc{Summer 2023}}
\makeoddfoot{SD}{}{\thepage}{}
\makeheadrule{SD}{\textwidth}{\normalrulethickness}
% \makefootrule{SD}{\textwidth}{\normalrulethickness}
\pagestyle{SD}

% Paragraph Indent
\nonzeroparskip
\setlength{\parindent}{0pt}

% Command to make writing sets easier
\newcommand{\set}[1]{\{\, #1\, \}}

% Absolute Value Operator
\DeclarePairedDelimiter\abs{\lvert}{\rvert}

\begin{document}
% -------------
\begin{center}
\LARGE{\textsc{Problem Set 2}}
 
\large{\textsc{Ms. Asma Sanam Larik}}
\end{center}

\begin{enumerate}
    % 1.
    \item
        Let \(\Sigma = \set{a, b, c}\) and \(L = \set{ {a^i}{b^j}{c^k} \mid i, j, k \ge 0 \text{ and if } i = 1,  \text{ then } j = k}\). Show that \textit{L} is not regular using the pumping lemma.
        \\\textbf{Hint:} You must show both cases where when \(w = xyz\) that \(x = \lambda\) and where \(x\) is composed of some characters.

        \textbf{Solution:}
        \\Let \(w = xyz\) and \(w = ab^pc^p\).
        \\If \(x = \lambda, y = ab^m\) and \(z = b^{p - m}c^p\),
        \\then \(w_i = \lambda(a^m)^i(b^{p - m}c^p)\).
        \\Pump up to get \(w_2 = (ab^m)^2(b^{p - m}c^p)
        \\= ab^mab^mb^{p - m}c^p
        \\= ab^mab^pc^p\)
        \\which does not belong to the language so we have a contradiction.
        \\And if \(x = ab^n, y = b^m\) and \(z = b^{p - m - n}c^p\),
        \\then \(w_i = (ab^n)(b^m)^i(b^{p - m - n}c^p)\).
        \\Pump up to get \(w_2 = (ab^n)(b^m)^2(b^{p - m - n}c^p)
        \\= ab^nb^mb^mb^{p - m - n}c^p
        \\= ab^{p + m}c^p\)
        \\which does not belong to the language so we have a contradiction.

        \textbf{Grading Scheme:}
        \\+1 mark - Chose a string \(w\) that belongs to the language
        \\+1 mark - Decompose \(w\) into \(w = xyz\) without 'fixing' any constraints
        \\+1 mark - Pump up/down \(w_i = xy_iz\)
        \\+1 mark - Show valid contradiction for \(x = \lambda\)
        \\+1 mark - Show valid contradiction for \(x \neq \lambda\)
        
    % 2.
    \item 
        Let \(\Sigma = \set{0, 1, +, =}\) and \(ADD = \{ u+v=w \mid u, v, w \in \set{0, 1}^* \text{ and } u+v=w \text{ is a valid addition operation}\}\). Show that \textit{ADD} is not regular using the pumping lemma.
        
        \textbf{Solution:}
        \\\(w\) can be any string which is a valid addition, for example:
        \\Let \(w = xyz\) and \(w => {1^p+0^p=1^p}\).
        \\And let \(x = 1^j, y = 1^k\) where \(|xy| \leq p\), and \(z => {1^{p - j - k}+0^p=1^p}\).
        \\Then \(w_i => {(1^j)(1^k)^i(1^{p - j - k}+0^p=1^p)}\).
        \\Pump up to get \(w_2 => {(1^j)(1^k)^2(1^{p - j - k}+0^p=1^p)}
        \\=> {1^j1^k1^k1^{p - j - k}+0^p=1^p}
        \\=> {1^{p+k}+0^p=1^p}\)
        \\which is not a valid addition operation so it does not belong to the language so we have a contradiction.

        \textbf{Grading Scheme:}
        \\+1 mark - Chose a string \(w\) that belongs to the language
        \\+1 mark - Decompose \(w\) into \(w = xyz\) without 'fixing' any constraints
        \\+1 mark - Pump up/down \(w_i = xy_iz\)
        \\+2 mark - Show valid contradiction
        
    % 3.
    \item
        For each of the following grammars, determine if they are ambiguous or unambiguous. If they are ambiguous, prove that by giving two parse trees for one same string.
        \begin{enumerate}
            % a)
            \item
                \(S \rightarrow XY \mid W \)
                \\\(X \rightarrow aXb \mid \lambda\)
                \\\(Y \rightarrow cY \mid \lambda\)
                \\\(W \rightarrow aWc \mid Z\)
                \\\(Z \rightarrow bZ \mid \lambda\)
            % b)
            \item
                \(S \rightarrow XX \)
                \\\(X \rightarrow aXb\)
                \\\(Y \rightarrow cY \mid \lambda\)
            % c)
            \item
                \(S \rightarrow aXY \mid bYX \mid \lambda \)
                \\\(Z \rightarrow aZ \mid a\)
                \\\(X \rightarrow aXY \mid a\)
                \\\(Y \rightarrow bYZ \mid b \mid \lambda\) 
        \end{enumerate}

        \textbf{Solution:}
        \\There was a correction made in the question in \textbf{(b)}, where the accurate grammar was supposed to be the following (and was communicated to all students beforehand):
        \\\textbf{\(S \rightarrow XX \)}
        \\\textbf{\(X \rightarrow aYb\)}
        \\\textbf{\(Y \rightarrow cY \mid \lambda\)}
        \\\textbf{(a)} \(\lambda\) can be generated in at least two different ways.
        \\\textbf{(b)} Any string of the form \(a^*b^*c^*\) can be generated in at least two different ways.
        \\\textbf{(c)} As one example, any string generated using the \(bYX\) rule has at least two parse trees.
        

        \textbf{Grading Scheme:}
        \\+0.5 marks - (a) is ambiguous
        \\+0.5 marks - (b) is ambiguous
        \\+0.5 marks - (c) is ambiguous
        \\+1 mark - Valid parse tree for (a)
        \\+1 mark - Valid parse tree for (b)
        \\+1 mark - Valid parse tree for (c)
        
    % 4.
    \item
        Construct a CFG for the following language over \(\Sigma = \set{\texttt{a}, \texttt{b}, \texttt{c}}\):
        \begin{enumerate}
            % a)
            \item
                \(A = \set{xby \mid x, y \in \{a\}^*}\) 
                \\\textbf{Solution:}
                \\Following or equivalent grammars are valid.
                \\\(S \rightarrow XbX\)
                \\\(X \rightarrow aX \mid \lambda\)
            % b)
            \item
                \(B = \set{ {a^i}{b^j}{c^k} \mid j = i + k}\)
                \\\textbf{Solution:}
                \\Following or equivalent grammars are valid.
                \\\(S \rightarrow AC\)
                \\\(A \rightarrow aAb \mid \lambda\)
                \\\(C \rightarrow bCc \mid \lambda\)
            % c)
            \item
                \(C = \set{ {a^i}{b^j}{c^k} \mid j = i \text{ or } j = k}\)
                \\\textbf{Solution:}
                \\Following or equivalent grammars are valid.
                \\\(S \rightarrow IC \mid AK\)
                \\\(I \rightarrow aIb \mid \lambda\)
                \\\(C \rightarrow cC \mid \lambda\)
                \\\(A \rightarrow aA \mid \lambda\)
                \\\(K \rightarrow bKc \mid \lambda\)
        \end{enumerate}

        \textbf{Grading Scheme:}
        \\+0.5 marks - Attempt at making any regular expression for
        
    % 5.
    \item
        Construct push-down automata for the following languages:
        \begin{enumerate}
            % a)
            \item
                \(A = \set{w \mid w \text{ is a palindrome i.e. } w = w^R \text{ and } w, w^R \in \{0, 1\}^*}\) 
            % b)
            \item
                \(B = \set{ {0^i}{1^j}{2^k} \mid i = j + k}\)
        \end{enumerate}

        \textbf{Solution:}
        \\Demonstrated in tutorial

        \textbf{Grading Scheme:}
        \\+0.5 marks - Attempt at making any regular expression for
        
\end{enumerate}
% -------------
\end{document}
