\documentclass[11pt]{article}
\usepackage{hyperref}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{epsfig}
\usepackage{amsfonts}
\usepackage{braket}


\newcommand{\handout}[5]{
  \noindent
  \begin{center}
  \framebox{
    \vbox{
      \hbox to 5.78in { {\bf CSE 309 Theory of Automata} \hfill #2 }
      \vspace{4mm}
      \hbox to 5.78in { {\Large \hfill #5  \hfill} }
      \vspace{2mm}
      \hbox to 5.78in { {\em #3 \hfill #4} }
    }
  }
  \end{center}
  \vspace*{4mm}
}

\newcommand{\lecture}[4]{\handout{#1}{#2}{#3}{#4}{#1}}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{fact}[theorem]{Fact}
\newtheorem{assumption}[theorem]{Assumption}

\topmargin 0pt
\advance \topmargin by -\headheight
\advance \topmargin by -\headsep
\textheight 8.9in
\oddsidemargin 0pt
\evensidemargin \oddsidemargin
\marginparwidth 0.5in
\textwidth 6.5in

\parindent 0in
\parskip 1.5ex
%\renewcommand{\baselinestretch}{1.25}

\begin{document} 

\lecture{Problem Set 3}{\textit{Assigned: Friday, 18 March}}{Spring 2021}{Due: \textit{11:59 pm Sunday, 3 April}}

\centerline{{\Large To facilitate grading and timely feedback}}
\centerline{{\Large please note that all submissions are through \href{http://gradescope.com}{Gradescope}.}}
\centerline{{\Large Solve each problem on a new page.}}

\begin{enumerate}

\item Let $A$ and $B$ be two disjoint languages. We say that language $C$ is a \emph{barrier} between $A$ and $B$ if $A \subseteq C$ and $B \subseteq \overline{C}$. Show that any two disjoint co-Turing recognizable languages have a decidable barrier language between them.

\item Let $A$ be a Turing-recognizable language consisting of descriptions of Turing machines, i.e.,
\begin{equation*}
A = \{\langle M_1 \rangle, \langle M_2 \rangle, \ldots \}, 
\end{equation*}
where every $M_i$ is a decider. Prove that some decidable language $D$ is not decided by any decider $M_i$ whose description appears in $A$.

\item Consider the problem of determining whether a single-tape Turing machine ever writes a blank symbol over a non-blank symbol during the course of its computation
on any input string. Formulate this problem as a language, and show that it is undecidable.

\item Let $\Gamma = \{0,1,B\}$ be the tape alphabet for all TMs in this problem, where $B$ is the blank symbol. Define the Sheikh Chilly function $SC: \mathbb{N} \mapsto \mathbb{N}$ as follows. For each value of $k$, consider all $k$-state TMs that halt when started with a blank tape. Let $SC(k)$ be the maximum number of $1$s that remain on the tape among all of these machines. Show that $SC$ is not a computable function.

\item Is the following language $X$ regarding single-tape TMs $M$ decidable? Prove your answer.
\begin{equation*}
X = \{ \langle M,w \rangle \, | \,  M \textrm{ never modifies the portion of the tape that contains the input } w \}.
\end{equation*}

\end{enumerate}
\end{document}

