\documentclass[11pt]{article}
\usepackage{hyperref}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{epsfig}
\usepackage{amsfonts}
\usepackage{braket}


\newcommand{\handout}[5]{
  \noindent
  \begin{center}
  \framebox{
    \vbox{
      \hbox to 5.78in { {\bf CSE 309 Theory of Automata} \hfill #2 }
      \vspace{4mm}
      \hbox to 5.78in { {\Large \hfill #5  \hfill} }
      \vspace{2mm}
      \hbox to 5.78in { {\em #3 \hfill #4} }
    }
  }
  \end{center}
  \vspace*{4mm}
}

\newcommand{\lecture}[4]{\handout{#1}{#2}{#3}{#4}{#1}}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{fact}[theorem]{Fact}
\newtheorem{assumption}[theorem]{Assumption}

\topmargin 0pt
\advance \topmargin by -\headheight
\advance \topmargin by -\headsep
\textheight 8.9in
\oddsidemargin 0pt
\evensidemargin \oddsidemargin
\marginparwidth 0.5in
\textwidth 6.5in

\parindent 0in
\parskip 1.5ex
%\renewcommand{\baselinestretch}{1.25}

\begin{document} 

\lecture{Problem Set 3}{\textit{Assigned: Saturday, 25 March}}{Spring 2023}{Due: \textit{11:59 pm Sunday, 9 April}}

\centerline{{\Large To facilitate grading and timely feedback}}
\centerline{{\Large please note that all submissions are through \href{http://gradescope.com}{Gradescope}.}}
\centerline{{\Large Solve each problem on a new page.}}

\begin{enumerate}

\item Modify Turing's proof of the undecidability of the halting problem to show there is no Turing machine $P$ with the following two properties:
\begin{enumerate}
\item For all Turing machines $M$, if $M()$ accepts then $P (\langle M \rangle)$ accepts, while if $M()$ rejects then $P(\langle M \rangle)$ rejects.

\item $P (x)$ halts for every input $x$. (In other words, if $M ( )$ runs forever, then $P (\langle M \rangle)$ can either accept or reject, but it cannot run forever.)
\end{enumerate}

\item Let a tile be a $1$-by-$1$ square, each of whose four edges has a color (chosen
from some finite set of colors). Two tiles ``snap together'' along an edge,
if and only if that edge has the same color on both of the tiles. The tiles
are all aligned the same way, and cannot be rotated. Then given a finite
set of tiles $C$, say that C ``fills the $n$-square'', if it's possible to fill a square
of size $n$-by-$n$ with $n^2$ tiles from $C$ (where each tile in $C$ can be duplicated
an unlimited number of times), so that every two neighboring tiles snap
together.

Call a set of tiles $C$ good if $C$ fills the $n$-square for every positive integer
$n$. Let $TILE$ be the problem of deciding whether $C$ is good, given a
description of $C$ as input. Show that $TILE$ is Turing-reducible to $HALT$.

\item Let $H(n)$ be the number of $n$-state Turing machines that halt when run on a blank tape. Also, let $COUNT$ be the problem of calculating $H(n)$, given a positive integer $n$ as input.
\begin{enumerate}
\item Show $COUNT$ is Turing-reducible to $HALT$.

\item Show $HALT$ is Turing-reducible to $COUNT$.

\item Let $L = \{\langle n, a \rangle \, | \, H(n) \geq a \}$. Show $L$ is Turing-recognizable.
\end{enumerate}

\item Prove that there does not exist a DTM $M$ that simultaneously satisfies both of the following two properties:
\begin{enumerate}
\item If $K$ is a DTM that halts on all input strings over its alphabet and $L(K)$ is infinite, then $M$ accepts $\langle K \rangle$.
\item If $K$ is a DTM that halts on all input strings over its alphabet and $L(K)$ is finite, then M does not accept $\langle K \rangle$.
\end{enumerate}
\emph{Hint:} Show that this implies that the language $ DIAG $ is Turing-recognizable, where
\begin{equation*}
DIAG = \{\langle T \rangle \, | \, T \textrm{ is a DTM and } \langle T \rangle \notin L(T) \}.
\end{equation*}

\end{enumerate}
\end{document}